Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-04-14 fib2.tex
mostly edits of the intro, some minor changes in the main text. references still missing though

\begin{abstract}
Classically simulating the dynamics of excitations in two-dimensional quantum systems that 
support anyons is likely intractable in general because such dynamics are sufficient for 
universal quantum computation. However, for typical processes of interest for the the study of 
quantum error correction in anyon systems, the relevant dynamics can in fact be efficiently 
simulated classically for a wide range of system parameters, even when the underlying anyon 
model is universal~\cite{RGsim}. We make use of this fact to demonstrate the success of an 
error-correction protocol for a Fibonacci anyon quantum memory. We numerically simulate a 
phenomenological model of the system and noise processes on lattice sizes of up to 
$128\times128$ sites, and find a lower bound on the error-correction threshold of 
approximately $12.5\%$, which is comparable to those previously known for abelian and 
(non-universal) nonabelian anyon models.
\end{abstract}

@@

Anyonic excitations in two-dimensional quantum systems~\cite{Wilczek1990} offer 
tremendous promise for long-term storage and processing of quantum 
information~\cite{Kitaev2003}. The topological features of quantum systems that support 
such excitations are insensitive to local 
perturbations~\cite{Bravyi2010, *Bravyi2011a, *Michalakis2013}, and the excitations 
themselves can in general be used to perform universal quantum 
computation~\cite{Freedman2002, *Freedman2002b}.

Quantum error correction is vital to harnessing the computational power of these topologically 
ordered systems. When coupled to a heat bath at any finite temperature, thermal fluctuations 
will create anyons that diffuse and quickly corrupt the stored quantum 
information~\cite{Pastawski2010}. Thus, the passive protection provided by the mass gap 
and low temperature must be augmented by an \emph{active} procedure.

Almost all of the sizable research effort on active quantum error correction for topological 
systems has focused on the case of abelian anyons~\cite{Terhal2014}. These systems are 
well suited to studying quantum error correction because the dynamics of excitations in 
abelian anyon models can be simulated numerically very efficiently, allowing lattice 
simulations of decoding with over 1 million sites~\cite{Duclos-Cianci2010}. The ability to 
efficiently simulate these systems is a consequence of the limitation that abelian anyons 
cannot be used for universal quantum computation.

Recent investigations have begun to explore quantum error correction for nonabelian anyon 
models~\cite{Brell2013, Wootton2013, Hutter2014}. These models are particularly interesting 
because braiding and fusion of anyons in these models leads in general to universal quantum 
computation. However, the initial studies have focused on models such as Ising 
anyons~\cite{Brell2013} and the so-called $\Phi$-$\Lambda$ 
model~\cite{Wootton2013, Hutter2014} that, while nonabelian, are not universal, a fact 
that was exploited to enable efficient classical simulation of error correction in these systems. 
The remarkable computational power of universal anyon models becomes a liability when we 
seek to simulate these systems on classical computers.

@@

\stf{Describe main results and why they are badass. cite:~\cite{Bonesteel2012}}

@@

Our error-correction algorithm is based on the hard-decision renormalization group (RG) 
decoders studied in Refs.~\cite{?}. Our decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, forming clusters of nearby anyons and fusing the anyons 
within each cluster, before iterating at increasing length scales. It terminates when the lattice 
is free of anyons (i.e.~we have returned to the codespace) or the clustering length scale 
exceeds the total size of the lattice. \cggb{A few more details? What is our length schedule?} 
\cggb{How does our decoder relate to other decoding schemes used in~\cite{Wootton2013,Brell2013,Hutter2014}?} 
Note that we expect that our qualitative results may be reproduced by most alternative 
families of decoders, such as matching decoders~\cite{?} and soft-decisions 
RG decoders~\cite{?} \cggb{do we?}. The advantage of using a hard-decision RG decoder is 
its simplicity and flexibility, and the fact that this simplicity allows for a proof of efficient 
classical simulability of successful error-correction~\cite{RGsim}.

@@

Although simulating pair creation, braiding, and fusion of $n$ Fibonacci anyons is equivalent 
in computational power to universal quantum computing (and thus unlikely to be classically 
tractable), noise processes and error-correction procedures have structure that we can 
exploit to efficiently simulate typical processes of interest~\cite{RGsim}. In particular, those 
processes in which we expect error-correction to succeed are also those that we expect to 
be able to efficiently simulate for the following heuristic reasons (see Ref.~\cite{RGsim} 
for a more rigorous analysis).

@@

\cggb {People we chatted to.} This work was supported by the ARC via 
EQuS project number CE11001013, and by the US Army Research Office grant numbers 
W911NF-14-1-0098 and W911NF-14-1-0103. STF also acknowledges support from an 
ARC Future Fellowship FT130101744. \cggb{CGB funding.}

>>>>

\begin{abstract}
Classically simulating the dynamics of excitations in two-dimensional quantum systems that 
support anyons is likely intractable in general because such dynamics are sufficient for 
universal quantum computation. However, processes of interest for the study of quantum error correction in anyon systems are typically drawn from a restricted class that displays significant structure over a wide range of system parameters.
We exploit this structure to classically simulate, and thus demonstrate the success of, an error-correction protocol for a quantum memory based on a universal anyon model, the Fibonacci anyons.  We numerically simulate a phenomenological model of the system and noise processes on lattice sizes of up to 
$128\times128$ sites, and find a lower bound on the error-correction threshold of 
approximately $12.5\%$, which is comparable to those previously known for abelian and 
(non-universal) nonabelian anyon models.
\end{abstract}

@@

Topologically ordered quantum systems in two dimensions offer tremendous promise for long-term storage and processing of quantum 
information~\cite{Kitaev2003}. The topological features of such systems are insensitive to local perturbations~\cite{Bravyi2010, Bravyi2011a, Michalakis2013}, and have excitations displaying anyonic statistics~\cite{Wilczek1990}. These anyons can in general be used as quantum memories or to perform universal topological quantum computation~\cite{Freedman2002, Freedman2002b, Nayak????}.

Quantum error correction is vital to harnessing the computational power of these topologically 
ordered systems. When coupled to a heat bath at any finite temperature, thermal fluctuations 
will create spurious anyons that diffuse and quickly corrupt the stored quantum 
information~\cite{Pastawski2010}. Thus, the passive protection provided by the mass gap 
at low temperature must be augmented by an \emph{active} procedure.

Almost all of the sizable research effort on active quantum error correction for topological 
systems has focused on the case of abelian anyons~\cite{Terhal2014}. These systems are 
well suited to studying quantum error correction because the dynamics of excitations in 
abelian anyon models (at the RG fixed point) can be simulated numerically very efficiently, allowing lattice 
simulations of decoding with over 1 million sites~\cite{Duclos-Cianci2010}. Standard algorithms to efficiently simulate these systems rely on the abelian nature of these particles, and in particular the fact that abelian anyons cannot be used for quantum computation.

Recent investigations have begun to explore quantum error correction for nonabelian anyon 
models~\cite{Brell2013, Wootton2013, Hutter2014}. Nonabelian anyon models are particularly interesting 
because braiding and fusion of these anyons in general allows for the implementation of universal quantum 
computation. However, the initial studies of error-correction in nonabelian anyon systems have focused on specific models, such as the Ising 
anyons~\cite{Brell2013} and the so-called $\Phi$-$\Lambda$ 
model~\cite{Wootton2013, Hutter2014} that, while nonabelian, are not universal. The general dynamics of these particular anyon models is known to be classically simulable, a fact
that was exploited to enable efficient simulation of error correction in these systems. When considering more general anyon models, their ability to perform universal quantum computation would seem a significant barrier to their simulation on a classical computer. While simulation of general dynamics does indeed seem intractable, we argue that the kinds of processes that are typical of thermal noise are sufficiently structured to allow for their classical simulation in precisely the regimes where we expect successful error correction to be possible. This insight allows us to simulate both noise and error-correction for a quantum code based on a universal anyon model.

@@

\stf{cite:~\cite{Bonesteel2012}} \cggb{what is this citation for? something about experimentally relevant in engineered systems?}

We use a flexible phenomenological model of dynamics and thermal noise to describe a system with Fibonacci anyon excitations. Within this model, we apply existing general topological error-correction protocols to simulate the successful preservation of quantum information encoded in topological degrees of freedom. Topological quantum computation protocols using nonabelian anyons typically implicitly assume the existence of an error-correction protocol to correct for diffusion or unwanted creation of anyons. Our results are the first explicit demonstration that such a scheme will be successful when applied to a universal topological quantum computer.

@@

Our error-correction algorithm is based on the hard-decision renormalization group (RG) 
decoders studied in Refs.~\cite{?}. Our decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, forming clusters of nearby anyons and fusing the anyons 
within each cluster, before iterating at increasing length scales. It terminates when the lattice 
is free of anyons (i.e.~we have returned to the codespace) or the clustering length scale 
exceeds the total size of the lattice. \cggb{A few more details? What is our length schedule?} 
\cggb{How does our decoder relate to other decoding schemes used in~\cite{Wootton2013,Brell2013,Hutter2014}?} 
Note that we expect that our qualitative results may be reproduced by most alternative 
families of decoders, such as matching decoders~\cite{?} and soft-decisions 
RG decoders~\cite{?} \cggb{do we?}. The advantage of using a hard-decision RG decoder is 
its simplicity and flexibility, and the fact that its clustering scheme does is compatible with the structure in the noise processes that allows us to classically simulate them.

@@

Although simulating pair creation, braiding, and fusion of $n$ Fibonacci anyons is equivalent 
in computational power to universal quantum computing (and thus unlikely to be classically 
tractable), noise processes and error-correction procedures have structure that we can 
exploit to efficiently simulate typical processes of interest. In particular, those 
processes in which we expect error-correction to succeed are also those that we expect to 
be able to efficiently simulate for the following heuristic reasons (we leave a more rigorous analysis of simulability for noise and error correction processes as an open problem).

@@

\cggb {People we chatted to.} This work was supported by the ARC via EQuS project number CE11001013, by the US Army Research Office grant numbers W911NF-14-1-0098 and W911NF-14-1-0103, ERC grant QFTCMPS and by the cluster of excellence EXC 201 Quantum Engineering and Space-Time Research. STF also acknowledges support from an ARC Future Fellowship FT130101744.
